A simple and efficient algorithm for computing approximate Nash equilibria in two-person zero-sum sequential games with imcomplete information

### 2015-06-25 nips2015.tex
refactor .tex
fixup algo label
fixups + more

\begin{algorithm}
\label{Tab:algo}
\caption{Primal-dual algorithm for computing approximate Nash
  Equilbria in two-person zero-sum games with imcomplete information
  and perfect recall}
\begin{algorithmic}[1]
\Require $\rho >0$, $\epsilon > 0$; $(y^{(0)},p^{(0)},x^{(0)},q^{(0)})
\in \mathbb{R}^{n_2}
  \times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1} \times
  \mathbb{R}^{l_2}$.
\Ensure A Nash $(\rho, \epsilon)$-equilibrium
$({y^*},{p^*},{x^*},{q^*}) \in S_1 \times S_2$ for
the GSP \eqref{eq:unconstrained_pb}.
\State  $\lambda \leftarrow 1/\|K\|$, ${v}_a^{(0)} \leftarrow 0$, $k
\leftarrow 0$
\While{$\|\tilde{\epsilon}_a^{(k)}\| \ge \epsilon$ or
  $\|{v}_{a}^{(k)}\|/k \ge \rho$}
\State $y^{(k + 1)} \leftarrow (y^{(k)} - \lambda (A^Tx^{(k)} +
E_2^Tq^{(k)}))_+$
\State  $p^{(k+1)} \leftarrow p^{(k)} -
\lambda(e_1-E_1x^{(k)})$
\State $x^{(k + 1)} \leftarrow (x^{(k)} + \lambda (Ay^{(k+1)} -
E_1^Tp^{(k+1)}))_+$, \hspace{.5em}$\Delta x^{(k+1)} \leftarrow
x^{(k+1)}-x^{(k)}$
\State $\Delta q^{(k+1)} \leftarrow
q^{(k+1)}-q^{(k)}$, \hspace{.5em}$q^{(k+1)}
\leftarrow q^{(k)} + \Delta q^{(k+1)}$
\State $y^{(k+1)} \leftarrow y^{(k+1)} - \lambda (A^T\Delta x^{(k+1)}
+ E_2^T\Delta q^{(k+1)})$, \hspace{.5em}$\Delta y^{(k+1)} \leftarrow
y^{(k+1)}-y^{(k)}$
\State $p^{(k+1)} \leftarrow p^{(k+1)} + \lambda E_1\Delta x^{(k+1)}$,
\hspace{.5em} $\Delta p^{(k+1)} \leftarrow p^{(k+1)}-p^{(k)}$
\State ${v}_a^{(k+1)} \leftarrow {v}_a^{(k)} +
\frac{1}{\lambda}(\Delta
y^{(k+1)},\Delta p^{(k+1)},\Delta x^{(k+1)},\Delta q^{(k+1)})$
\State $k \leftarrow k + 1$
\EndWhile
\end{algorithmic}
\end{algorithm}

@@

\begin{proof}
Note that the quadruplet $(\Psi_1, \Psi_2, g_1, g_2)$ clearly satisfies
assumptions B.1, B.2, B.3, B.5, and B.6 of \cite{he2013accelerating}
with $L_{xx} = L_{yy} = 0$ and $L_{xy} = L_{yx} = \|K\|$. Now,
one easily computes the proximal operators $g_1$ and $g_2$ as
\begin{eqnarray}
  \prox_{\lambda g_1}(y, p) \equiv ((y)_+, p - \lambda
  e_1),\hspace{.5em} \prox_{\lambda g_2}(x, q) \equiv ((x)_+, q -
  \lambda e_2).
  \end{eqnarray}

@@

$E_1 \in \mathbb{R}^{3 \times 5}$ with $E_1(0,0) = E_1(1,1) =
E_1(1,2) = E_1(2,3) = E_1(2,4) = \textbf{1}$,
$E_1(1,0) = E_1(2,0) = \textbf{-1}$; \\

$E_2 \in \mathbb{R}^{3 \times 5}$ with $E_2(0,0) = E_2(1,1) = E_2(1,2)
= E_2(2,3) = E_2(2,4) = \textbf{1}$, $E_2(1,0) = E_2(2,0) =
\textbf{-1}$; and \\

$A \in \mathbb{R}^{5 \times 5}$ with $A(2,0) =
A(4,0) = \textbf{-0.5}$, $A(1,3) = \textbf{1}$, $A(3,1) =
\textbf{-1}$, $A(1,2) = A(1,4) = A(3,2) = A(3,4) = \textbf{0.25}$.

\paragraph{\textbf{Kuhn 3-card Poker.}}
The game has $n_1 = n_2 = 13$ and $l_1 = l_2 = 7$, and its
sequence-form is given by:\\

$E_1 \in \mathbb{R}^{7 \times 13}$ with $E_1(0,0) = E_1(1,9) =
E_1(1,12) = E_1(2,1) = E_1(2,4) = E_1(3,5) =
E_1(3,8) = E_1(4,2) = E_1(4,3) = E_1(5,6) = E_1(5,7) = E_1(6,10) =
E_1(6,11) = \textbf{1}$, $E_1(1,0) = E_1(2,0) = E_1(3,0) = E_1(4,1) =
E_1(5,5) = E_1(6,9) = \textbf{-1}$; \\

 $E_2 \in \mathbb{R}^{7 \times 13}$
$E_2(0,0) = E_2(1,7) = E_2(1,8) = E_2(2,9) = E_2(2,10) = E_2(3,5) =
E_2(3,6) = E_2(4,11) = E_2(4,12) = E_2(5,1) = E_2(5,2) = E_2(6,3) =
E_2(6,4) = \textbf{1}$, $E_2(1,0) = E_2(2,0) = E_2(3,0) = E_2(4,0) =
E_2(5,0) = E_2(6,0) = \textbf{-1}$; and \\

 $A \in \mathbb{R}^{13 \times
  13}$ with $A(3,8) = A(3,12) = A(4,6) = A(4,10) = A(7,12) = A(8,10) =
\textbf{-0.333333}$, $A(1,7) = A(1,11) = A(2,8) = A(2,12) = A(5,11) =
A(6,4) = A(6,12) = A(10,4) = A(10,8) = \textbf{-0.166667}$, $A(7,4) =
A(8,2) = A(11,4) = A(11,8) = A(12,2) = A(12,6) = \textbf{0.333333}$,
$A(4,5) = A(4,9) = A(5,3) = A(8,1) = A(8,9) = A(9,3) = A(9,7) =
A(12,1) = A(12,5) = \textbf{0.166667}$.\\

@@

\begin{eqnarray*}
  \begin{split}
    E_2y^* - e_2 = [&-7.04 \times 10^{-7}, 2.27 \times 10^{-6}, -3.29
      \times 10^{-6}, -1.50 \times 10^{-6},\\
      &2.92 \times 10^{-6}, -4.97 \times 10^{-7}, -5.85 \times
      10^{-7}]^T
    \end{split}
\end{eqnarray*}
Finally, one checks that
\begin{eqnarray*}
  {x^*}^TAy^* = \textbf{-0.05555},
\end{eqnarray*}
 which agrees to 5 d.p with the value of $-1 / 18$ computed
 analytically by H. W. Kuhn in his 1950 paper \cite{kuhn}. The
 evolution of the dual gap and the expected value of the game across
 iterations are shown in Figure \ref{Tab:dgap_curve}.

\section{Conclusion}
Making use of the sequence-form representation
\cite{koller1992complexity,von1996efficient,vonequilibrium}, we have
deviced a primal-dual algorithm for computing Nash equilibria in
two-person zero-sum sequential games with imcomplete information (like
Texas Hold'em, etc.). Our algorithm is simple to implement, with a
very low constant cost per iteration, and enjoys a rigorous
convergence theory with a proven $\mathcal{O}(1/min(\rho,\epsilon))$ convergence
in terms of basic operations (matvec products, clipping, etc.), to a
Nash $(\rho,\epsilon$-equilibrium of the game.

Equilibrium problems are saddle-point convex-concave problems, and as
such a natural choice for an algorithm would be in the family of
primal-dual algorithms. The author believes such primal-dual schemes
will receive more attention in the algorithmic game theory community
in future.

>>>>

\begin{algorithm}
\caption{Primal-dual algorithm for computing approximate Nash
  Equilbria in two-person zero-sum games with imcomplete information
  and perfect recall}
\label{Tab:algo}
\begin{algorithmic}[1]
\Require $\rho >0$, $\epsilon > 0$; $(y^{(0)},p^{(0)},x^{(0)},q^{(0)})
\in \mathbb{R}^{n_2}
  \times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1} \times
  \mathbb{R}^{l_2}$.
\Ensure A Nash $(\rho, \epsilon)$-equilibrium
$({y^*},{p^*},{x^*},{q^*}) \in S_1 \times S_2$ for
the GSP \eqref{eq:unconstrained_pb}.
\State  $\lambda \leftarrow 1/\|K\|$, ${v}_a^{(0)} \leftarrow 0$, $k
\leftarrow 0$
\While{$\|\tilde{\epsilon}_a^{(k)}\| \ge \epsilon$ or
  $\|{v}_{a}^{(k)}\|/k \ge \rho$}
\State $y^{(k + 1)} \leftarrow (y^{(k)} - \lambda (A^Tx^{(k)} +
E_2^Tq^{(k)}))_+$
\State  $p^{(k+1)} \leftarrow p^{(k)} -
\lambda(e_1-E_1x^{(k)})$
\State $x^{(k + 1)} \leftarrow (x^{(k)} + \lambda (Ay^{(k+1)} -
E_1^Tp^{(k+1)}))_+$, \hspace{.5em}$\Delta x^{(k+1)} \leftarrow
x^{(k+1)}-x^{(k)}$
\State $\Delta q^{(k+1)} \leftarrow
q^{(k+1)}-q^{(k)}$, \hspace{.5em}$q^{(k+1)}
\leftarrow q^{(k)} + \Delta q^{(k+1)}$
\State $y^{(k+1)} \leftarrow y^{(k+1)} - \lambda (A^T\Delta x^{(k+1)}
+ E_2^T\Delta q^{(k+1)})$, \hspace{.5em}$\Delta y^{(k+1)} \leftarrow
y^{(k+1)}-y^{(k)}$
\State $p^{(k+1)} \leftarrow p^{(k+1)} + \lambda E_1\Delta x^{(k+1)}$,
\hspace{.5em} $\Delta p^{(k+1)} \leftarrow p^{(k+1)}-p^{(k)}$
\State ${v}_a^{(k+1)} \leftarrow {v}_a^{(k)} +
\frac{1}{\lambda}(\Delta
y^{(k+1)},\Delta p^{(k+1)},\Delta x^{(k+1)},\Delta q^{(k+1)})$
\State $k \leftarrow k + 1$
\EndWhile
\end{algorithmic}
\end{algorithm}

@@

\begin{proof}
It is clear to see that the quadruplet $(\Psi_1, \Psi_2, g_1, g_2)$
satisfies assumptions B.1, B.2, B.3, B.5, and B.6 of
\cite{he2013accelerating} with $L_{xx} = L_{yy} = 0$ and $L_{xy} =
L_{yx} = \|K\|$. Now, one easily computes the proximal operators $g_1$
and $g_2$ as
\begin{eqnarray*}
  \prox_{\lambda g_1}(y, p) \equiv ((y)_+, p - \lambda
  e_1),\hspace{.5em} \prox_{\lambda g_2}(x, q) \equiv ((x)_+, q -
  \lambda e_2).
  \end{eqnarray*}

@@

$E_1 \in \mathbb{R}^{3 \times 5}$ with $E_1(0,0) = E_1(1,1) =
E_1(1,2) = E_1(2,3) = E_1(2,4) = {1}$,
$E_1(1,0) = E_1(2,0) = {-1}$; \\

$E_2 \in \mathbb{R}^{3 \times 5}$ with $E_2(0,0) = E_2(1,1) = E_2(1,2)
= E_2(2,3) = E_2(2,4) = {1}$, $E_2(1,0) = E_2(2,0) =
{-1}$; and \\

$A \in \mathbb{R}^{5 \times 5}$ with $A(2,0) =
A(4,0) = {-0.5}$, $A(1,3) = {1}$, $A(3,1) =
{-1}$, $A(1,2) = A(1,4) = A(3,2) = A(3,4) = {0.25}$.

\paragraph{{Kuhn 3-card Poker.}}
The game has $n_1 = n_2 = 13$ and $l_1 = l_2 = 7$, and its
sequence-form is given by:\\

$E_1 \in \mathbb{R}^{7 \times 13}$ with $E_1(0,0) = E_1(1,9) =
E_1(1,12) = E_1(2,1) = E_1(2,4) = E_1(3,5) =
E_1(3,8) = E_1(4,2) = E_1(4,3) = E_1(5,6) = E_1(5,7) = E_1(6,10) =
E_1(6,11) = {1}$, $E_1(1,0) = E_1(2,0) = E_1(3,0) = E_1(4,1) =
E_1(5,5) = E_1(6,9) = {-1}$; \\

 $E_2 \in \mathbb{R}^{7 \times 13}$ with
$E_2(0,0) = E_2(1,7) = E_2(1,8) = E_2(2,9) = E_2(2,10) = E_2(3,5) =
E_2(3,6) = E_2(4,11) = E_2(4,12) = E_2(5,1) = E_2(5,2) = E_2(6,3) =
E_2(6,4) = {1}$, $E_2(1,0) = E_2(2,0) = E_2(3,0) = E_2(4,0) =
E_2(5,0) = E_2(6,0) = {-1}$; and \\

 $A \in \mathbb{R}^{13 \times
  13}$ with $A(3,8) = A(3,12) = A(4,6) = A(4,10) = A(7,12) = A(8,10) =
{-0.333333}$, $A(1,7) = A(1,11) = A(2,8) = A(2,12) = A(5,11) =
A(6,4) = A(6,12) = A(10,4) = A(10,8) = {-0.166667}$, $A(7,4) =
A(8,2) = A(11,4) = A(11,8) = A(12,2) = A(12,6) = {0.333333}$,
$A(4,5) = A(4,9) = A(5,3) = A(8,1) = A(8,9) = A(9,3) = A(9,7) =
A(12,1) = A(12,5) = {0.166667}$.\\

@@

\begin{eqnarray*}
  \begin{split}
    E_2y^* - e_2 = [&-7.04 \times 10^{-7}, 2.27 \times 10^{-6}, -3.29
      \times 10^{-6}, -1.50 \times 10^{-6},\\
      &2.92 \times 10^{-6}, -4.97 \times 10^{-7}, -5.85 \times
      10^{-7}]^T
    \end{split}
\end{eqnarray*}
Finally, one checks that
\begin{eqnarray*}
  {x^*}^TAy^* = {-0.05555},
\end{eqnarray*}
 which agrees to 5 d.p with the value of $-1 / 18$ computed
 analytically by H. W. Kuhn in his 1950 paper \cite{kuhn}. The
 evolution of the dual gap and the expected value of the game across
 iterations are shown in Figure \ref{Tab:dgap_curve}.

\section{Conclusion}
Making use of the sequence-form representation
\cite{koller1992complexity,von1996efficient,vonequilibrium}, we have
deviced a primal-dual algorithm for computing Nash equilibria in
two-person zero-sum sequential games with imcomplete information (like
Texas Hold'em, etc.). Our algorithm is simple to implement, with a
very low constant cost per iteration, and enjoys a rigorous
convergence theory with a proven $\mathcal{O}(1/min(\rho,\epsilon))$ convergence
in terms of basic operations (matvec products, clipping, etc.), to a
Nash $(\rho,\epsilon$-equilibrium of the game.
Equilibrium problems are saddle-point convex-concave problems, and as
such a natural choice for an algorithm would be in the family of
primal-dual algorithms. The author believes such primal-dual schemes
will receive more attention in the algorithmic game theory community
in future.
